---
title: "Raminski UCB R Class Asgt 2"
author: "Michael Raminski"
date: "Friday, March 31, 2017"
output: html_document
---

##Diary

For this assignment, I had trouble with a few items:
-At first, I could not run the "glimpse" function, but that problem was solved by downloading the "plyr" and "dplyr" packages
-Then, I could only run the "glimpse" function in an outside console, I couldn't run my markdown file with the glimpse function- it always resulted in an error. However, with the help of peers, I realized that I needed to reload the packages in my R Markdown code
-This was also the case with the "write_csv" function. I could run that function in an outside console, but I could not run that function in the R Markdown file- it always results in an error. Again, I had to run the function to load the necessary package in order to get the file to knit
-I'm still struggling a bit with the syntax of the below function (although the function works):

    Import[Import$Year_of_Loss >= 1990,]

    -I don't understand why the "," is necessary after "1990"
    -I don't understand why I get an error if I write the function as such:
    
    Import["Year_of_Loss" >= 1990,]
    


##Summary of Results

Aside from the issues that I noted above, the results of the assignment were mostly as expected. I was even able to try several different approaches to some of the tasks and still got the same results. The one thing that I didn't understand was that when I ran the "write_csv" function, it produced a spreadsheet/dataframe, rather than a single column of comma seperated values. I assume the assignment was asking to reimport the data from a comma seperated value format back to a dataframe format. Therefore, I included the code to do that as well. 



##Assignment

###Task 2

####Import dataset

```{r, warning=FALSE, message=FALSE, error=FALSE, eval=TRUE, include=FALSE}
Import <- read.csv("Aviation.csv")
```
######Import <- read.csv("Aviation.csv")



####Exploring Dataset - Column Names

```{r, warning=FALSE, message=FALSE, error=FALSE, eval=TRUE}
colnames(Import)
```



####Exploring Dataset - Dimensions

```{r, warning=FALSE, message=FALSE, error=FALSE, eval=TRUE}
dim(Import)
```



####Exploring Dataset - Structure

```{r, warning=FALSE, message=FALSE, error=FALSE, eval=TRUE}
str(Import)
```



####Exploring Dataset - Top of dataset

```{r, warning=FALSE, message=FALSE, error=FALSE, eval=TRUE}
head(Import)
```



####Exploring Dataset - Bottom of dataset

```{r, warning=FALSE, message=FALSE, error=FALSE, eval=TRUE}
tail(Import)
```



####Exploring Dataset - Glimpse

```{r, warning=FALSE, message=FALSE, error=FALSE, eval=TRUE}
library(plyr)
library(dplyr)
glimpse(Import)
```



###Task 3

####Changing Dataset - Rename columns to 1 through 19

```{r, warning=FALSE, message=FALSE, error=FALSE, eval=TRUE, include=FALSE}
colnames(Import) <- c(1:19)
Import
```



####Changing Dataset - Remove columns, only use first 10 columns

```{r, warning=FALSE, message=FALSE, error=FALSE, eval=TRUE, include=FALSE}
Import[1:10]
```
######Import[1:10]
###### OR
######Import[-11:-19]



####Changing Dataset - Exclude rows by logical test, only show records with a "Year of Loss" of 1990 or after (show only header)

```{r, warning=FALSE, message=FALSE, error=FALSE, eval=TRUE}
Import <- read.csv("Aviation.csv")
head(Import[Import$Year_of_Loss >= 1990,])
```
###### OR
######Import[Import[2] >= 1990,]


###Task 4

####Factors - Change the "stringsasfactors" setting for the "Phase_of_Flight" column (show only the head of results)

```{r, warning=FALSE, message=FALSE, error=FALSE, eval=TRUE}
zz <- data.frame(Import[7], stringsAsFactors = TRUE)
head(zz)
```



####Factors - The data does not appear to be changed by changing the "stringsasfactors" setting

```{r, warning=FALSE, message=FALSE, error=FALSE, eval=TRUE}
zz <- data.frame(Import[7], stringsAsFactors = FALSE)
head(zz)
```



####Factors - Create factors for the "Year_of_Loss" column (show only the head of results)

```{r, warning=FALSE, message=FALSE, error=FALSE, eval=TRUE}
factored <- within(Import[2], {
  Howold <- NA
  Howold[Year_of_Loss < 1990] <- "Very Old"
  Howold[Year_of_Loss >= 1990 & Year_of_Loss <= 2000] <- "Kind of Old"
  Howold[Year_of_Loss > 2000] <- "New"
})
head(factored)
```



###Task 5

####Exporting Dataframe to .csv file 

```{r, warning=FALSE, message=FALSE, error=FALSE, eval=TRUE, include=FALSE}
library(readr)
write_csv(Import, "Import.csv")
```
######write_csv(Import, "Import.csv")



####Reimporting Dataframe

```{r, warning=FALSE, message=FALSE, error=FALSE, eval=TRUE, include=FALSE}
Reimport <- read.csv("Import.csv")
Reimport
```
######Reimport <- read.csv("Import.csv")
######Reimport
######OR
######If data is exported in single-column comma seperated value format (rather than data frame/spreadsheet format)
######read.table("Import.csv", header=TRUE, sep=";")
