---
title: "Raminski UCB R Class Asgt 1"
output: html_document
---

##Diary
I had some trouble starting the assignment because I was still grasping the concept of a "Markdown" file- format, purpose, how to build, etc.

Additionally, I downloaded MiKTeX, but had some trouble at the beginning generating a Markdown file in Word. Attempting to do so gave me an error message that said that MiKTeX was not installed. However, that seems to have been resolved somehow and now I do have the option of generating a Word Markdown file (although I will submit in html)



##Summary of Results
The results generated were largely as expected especially since I had read the materials before completing the assignment.

However, I was still not sure how to summarize the large "Arthritis" table. When I added "results=hide" to the code, I was given an error message. I am still curious about the correct way to either summarize or hide the results entirely.



##Kabacoff Section 1.8

1. Open the general help, and look at the "Introduction to R" section

```{r, warning=FALSE, message=FALSE, error=FALSE}
help.start()
```

2. Install the vcd package (code not run)

install.packages("vcd")

3. List the functions and datasets available in this package

```{r, warning=FALSE, message=FALSE, error=FALSE}
help(package="vcd")
```

4. Load the package, and read the description of dataset Arthritis

```{r, warning=FALSE, message=FALSE, error=FALSE}
library(vcd)
```

5. Print out the Arthritis dataset

```{r, warning=FALSE, message=FALSE, error=FALSE}
Arthritis
```

6. Run the example that comes with the Arthritis dataset

```{r, warning=FALSE, message=FALSE, error=FALSE}
example(Arthritis)
```

7. Quit (code not run)

q()


##Teetor Section 5.1-5.19
###Section 5.1

Appending a single item in a vector

```{r, warning=FALSE, message=FALSE, error=FALSE}
v <- c(1,2,3)
v <- c(v,4) # Append a single value to v
v
```

Appending multiple items in a vector

```{r, warning=FALSE, message=FALSE, error=FALSE}
w <- c(5,6,7,8)
v <- c(v,w) # Append an entire vector to v
v
```

Append an items by assigning it to a poistion in a vector

```{r, warning=FALSE, message=FALSE, error=FALSE}
v <- c(1,2,3) # Create a vector of three elements
v[10] <- 10 # Assign to the 10th element
v # R extends the vector automatically
```

###Section 5.2

Insert an item into a vector

```{r, warning=FALSE, message=FALSE, error=FALSE}
append(1:10, 99, after=5)
```

Insert an item into a vector

```{r, warning=FALSE, message=FALSE, error=FALSE}
append(1:10, 99, after=0)
```

###Section 5.3

The "recycling rule" - handling vectors of different lengths

```{r, warning=FALSE, message=FALSE, error=FALSE}
c (1:3)
c (1:6)
(1:6) + (1:3)
```

Use "cbind" to create column vectors

```{r, warning=FALSE, message=FALSE, error=FALSE}
cbind (1:3)
cbind (1:6)
cbind((1:6), (1:3))
```

If shorter vector is not a multiple of the larger vector then R generates a warning

```{r, message=FALSE, error=FALSE}
(1:5) + (1:6)
```

Add a scalar value to a vector

```{r, warning=FALSE, message=FALSE, error=FALSE}
(1:6) + 10
```
















